---
title: "Project 3"
author: "Anthony Ferraro, Lauren Moylan, Anna Weiss"
output:
  html_document:
    code_folding: hide
    df_print: paged
linkcolor: blue
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning= FALSE)

library(readxl)
library(readr)
library(ggplot2)
library(mosaic)
library(forcats)
library(dplyr)
```

### Introduction

We will be looking at football data from the years 2024 and 2025. Specifically, 
we will look at rushing data for these years. Originally, each data set had 501 
observations, but we will be filtering the data down to only the top 50 
observations in the yards per attempt variable (Y/A). Then, we will be looking 
at how yards per attempt is impacted by age.


### Data Import
```{r}
Data_24 <- read_csv("2024 Data Actual - Sheet1.csv") 

Data_25 <- read_csv("2025 Data - Sheet1.csv")

Data_24 = Data_24 %>%
  mutate(Season = 2024, as.factor(Season))

Data_25 = Data_25 %>%
  mutate(Season = 2025, as.factor(Season))
```

We gathered each data set from pro football reference and downloaded it as a 
CSV file. 

https://www.pro-football-reference.com/years/2024/rushing.htm
https://www.pro-football-reference.com/years/2025/rushing.htm


### Exploratory Data Analysis

#### Yards Per Attempt Variable
```{r}
knitr::kable(favstats(~`Y/A`, data = Data_24))

knitr::kable(favstats(~`Y/A`, data = Data_25))
```

We can see that there was a higher minimum yards per attempt in 2025 but a 
lower maximum yards per attempt. Also, there was a higher average yards per 
attempt among players in the 2024 season then the 2025 season.


#### Age Variable
```{r}
ggplot(Data_24, aes(x = Age)) +
  geom_bar(fill = "darkgreen") +
  theme_bw() +
  labs(
    title = "2024"
  )

ggplot(Data_25, aes(x = Age)) +
  geom_bar(fill = "darkgreen") +
  theme_bw() +
  labs(
    title = "2025"
  )
```

We can see in both 2024 and 2025 that there is a cluster of rushers with an age 
of 25, which is considered the prime age for football players so that isn't 
surprising. However, there seems to be a lot more rushers under the age on 25 
in 2025 than 2024.


### Research - How Age Impacts Yards per Attempt
```{r}
Combined_Data = bind_rows(Data_24, Data_25)

ggplot(Combined_Data, aes(x = Age, y = `Y/A`, color = Player, group = Player)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "none") 

```


### Conclusion
From this graph, we can see that it is hard to find a significant difference 
among rushing performance year over year for players under the age of 27. 
There are many different outcomes, with players increasing or decreasing their
production with no real pattern. However, it is possible that at age 27 or older,
rushers start to see more of a significant drop-off as out of the 9 rushers 
age 27 or older that we have multiple seasons of data for, 7 of them got worse 
or stayed the same.



